---
title: "Embeddings and Cluster Discovery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Embeddings and Cluster Discovery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
has_cluster_deps <- requireNamespace("dbscan", quietly = TRUE) &&
  requireNamespace("uwot", quietly = TRUE) &&
  requireNamespace("tidytext", quietly = TRUE) &&
  requireNamespace("ggrepel", quietly = TRUE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  eval = has_cluster_deps
)
```

## Semantic research landscape

scimapR uses text embeddings to discover research clusters and visualise
the intellectual structure of a field.

```{r setup}
library(scimapR)

# Example corpus comes with pre-computed embeddings
corpus <- sm_example_corpus(with_embeddings = TRUE, seed = 42)
```

## Clustering

### HDBSCAN

```{r hdbscan}
corpus <- sm_cluster_hdbscan(corpus, min_cluster_size = 10)
table(corpus$works$cluster_id)
```

### Visualise the landscape

```{r landscape, fig.cap = "Research landscape coloured by cluster"}
sm_plot_landscape(corpus, color_by = "cluster_id", reducer = "umap")
```

### Cluster labels (TF-IDF)

```{r labels}
labels <- sm_cluster_label(corpus, method = "tfidf", n_terms = 5)
head(labels)
```

## Thematic map

```{r thematic, fig.cap = "Strategic diagram"}
sm_plot_thematic_map(corpus)
```

## Advanced: Python-based embeddings

For production use, compute embeddings with SPECTER2:

```r
# Requires Python + sentence-transformers
corpus <- sm_embed_works(corpus, model = "specter2")
```

Alternatively, fetch pre-computed SPECTER embeddings from Semantic Scholar:

```r
corpus <- sm_enrich_specter(corpus)
```
