---
title: "Equity Audit, Trajectory Analysis, and Corpus Chat"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Equity Audit, Trajectory Analysis, and Corpus Chat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Equity and representation audit

scimapR includes built-in tools for auditing geographic, gender, funding,
and open-access representation -- with explicit confidence reporting and
limitation caveats.

```{r setup}
library(scimapR)
corpus <- sm_example_corpus(seed = 42)
```

### Geographic audit

```{r geo}
geo <- sm_audit_geographic(corpus)
head(geo)
```

### Open access audit

```{r oa}
oa <- sm_audit_oa(corpus)
head(oa)
```

### Equity dashboard

```{r dashboard, fig.width = 8, fig.height = 6, fig.cap = "Equity dashboard"}
sm_plot_equity_dashboard(corpus)
```

### Epistemic humility

All audit functions report confidence, method, and caveats. The print
methods include explicit limitation blocks. scimapR does not pretend
that automated inference is ground truth.

## Author trajectory analysis

scimapR can track an author's career evolution through topic pivots,
collaborator turnover, and productivity curves.

```{r trajectory}
# The example corpus includes a trajectory seed author
traj <- sm_author_trajectory(corpus, author_id = "A000000001")
print(traj)
```

```{r traj-plot, fig.width = 8, fig.height = 6, fig.cap = "Author trajectory"}
sm_plot_trajectory(traj)
```

## LLM-grounded corpus chat

scimapR can use an LLM to answer questions about a corpus, with every
claim anchored to actual works via retrieval-constrained citations.

```r
# Requires ellmer package
response <- sm_chat(
  corpus,
  "What methodological approaches dominate post-2020?",
  provider = ellmer::chat_anthropic()
)
print(response)
```

The LLM only sees works that were actually retrieved. Citations are
constrained to those retrieved works -- no hallucinated references are
possible.
