---
title: "Building a Corpus from Files and APIs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building a Corpus from Files and APIs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(scimapR)
```

## Bring your own relational data (recommended)

If your data is already tabular -- a manual tracker, an institutional export, a
spreadsheet of works -- the most robust entry point is
`sm_corpus_from_tables()`. It validates the required columns, coerces types,
fills missing optional tables with correctly-typed 0-row tibbles, and returns a
ready `sm_corpus`, side-stepping format-specific parsers entirely. A works-only
corpus is perfectly valid:

```{r from-tables}
works <- data.frame(
  work_id = paste0("W", 1:3),
  title = c("Spatial transcriptomics in cancer",
            "Immune checkpoint resistance",
            "A biomarker discovery cohort"),
  year = c("2019", "2020", "2021"),   # character -> coerced to integer
  doi = paste0("10.1234/example.", 1:3),
  cited_by_count = c(12, 5, 1)
)
corpus <- sm_corpus_from_tables(list(works = works))
corpus
```

You can pass any of `works`, `authorships`, `sources`, `institutions`,
`references`, `concepts`, etc. as named list elements.

## File ingestion

scimapR reads 12 bibliographic formats with native clean-room parsers.

```{r files}
# List available example files
sm_example_files()
```

```{r read-bib}
# Read a BibTeX file
bib_path <- sm_example_files("example.bib")
corpus_bib <- sm_read_bib(bib_path)
print(corpus_bib)
```

```{r read-ris}
# Read an RIS file
ris_path <- sm_example_files("example.ris")
corpus_ris <- sm_read_ris(ris_path)
nrow(corpus_ris$works)
```

## Combining multiple sources

```{r combine}
combined <- sm_build_corpus(corpus_bib, corpus_ris, dedupe = TRUE)
nrow(combined$works)
```

## Auto-detection

```{r auto}
corpus_auto <- sm_read_auto(bib_path)
nrow(corpus_auto$works)
```

## API fetching

For API-based ingestion, set your email for polite pool access:

```r
Sys.setenv(SCIMAPR_MAILTO = "your.email@example.com")

# Fetch from OpenAlex
corpus <- sm_fetch_openalex(
 query = "spatial transcriptomics colorectal cancer",
 n_max = 100
)
```

All API fetchers populate the provenance table for reproducibility tracking.

## Deduplication

```{r dedupe}
corpus <- sm_example_corpus()
deduped <- sm_dedupe(corpus)
```

## Validation

```{r validate}
issues <- sm_validate(corpus)
nrow(issues)
```
