---
title: "Modern Citation Indicators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modern Citation Indicators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/CTTIR/scimapR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CTTIR/scimapR/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/CTTIR/scimapR/actions/workflows/pkgdown.yaml/badge.svg)](https://cttir.github.io/scimapR/)
[![CRAN status](https://www.r-pkg.org/badges/version/scimapR)](https://CRAN.R-project.org/package=scimapR)
[![Codecov test coverage](https://codecov.io/gh/CTTIR/scimapR/branch/main/graph/badge.svg)](https://app.codecov.io/gh/CTTIR/scimapR?branch=main)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/scimapR)](https://cran.r-project.org/package=scimapR)
[![CRAN downloads total](https://cranlogs.r-pkg.org/badges/grand-total/scimapR)](https://cran.r-project.org/package=scimapR)
[![License: MIT](https://img.shields.io/badge/license-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Beyond the h-index

scimapR implements several modern citation indicators that go beyond
simple counts.

```{r setup}
library(scimapR)
corpus <- sm_example_corpus(seed = 42)
```

## Classical indices

```{r h-index}
h <- sm_metric_h_index(corpus, level = "author")
head(h, 10)
```

## CD index (disruption)

The CD index (Funk & Owen-Smith, 2017) measures whether a work
consolidates or disrupts its field.

```{r disruption}
cd <- sm_metric_disruption(corpus)
head(cd)
```

## Relative Citation Ratio

The RCR (Hutchins et al., 2016) normalises citations relative to the
co-citation network.

```{r rcr}
rcr <- sm_metric_rcr(corpus)
head(rcr)
```

## Field-Normalized Citation Impact

FNCI (Waltman et al., 2011) normalises by field and year.

```{r fnci}
fnci <- sm_metric_fnci(corpus)
head(fnci)
```

## Uzzi novelty

Measures atypical journal combinations in reference lists (Uzzi et al., 2013).

```{r novelty}
nov <- sm_metric_novelty(corpus)
head(nov)
```

## Summary tables

```{r summary}
sm_summary_authors(corpus) |> head(5)
```

## Self-citation and self-corrected indices

`sm_self_citation()` derives author (or institution) self-citation from the
reference lists already in the corpus --- no per-citation API calls. It returns
per-entity and per-work tibbles plus a provenance trail showing which works
drove each self-citation, suitable for an institutional report.

```{r self-citation}
sc_corpus <- readRDS(system.file("extdata", "example_self_citation_corpus.rds",
                                 package = "scimapR"))
sc <- sm_self_citation(sc_corpus, level = "author")
sc$by_entity
head(sc$provenance)
```

The h/g/m indices accept `self_corrected = TRUE`, which recomputes the index
after removing those self-citations (always `<=` the uncorrected value):

```{r self-corrected}
merge(
  sm_metric_h_index(sc_corpus, "author"),
  sm_metric_h_index(sc_corpus, "author", self_corrected = TRUE),
  by = "author_id", suffixes = c("", "_corrected")
)
```

## References

- Funk, R. J. & Owen-Smith, J. (2017). A Dynamic Network Measure of
  Technological Change. *Management Science*, 63(3), 791--817.
- Hutchins, B. I. et al. (2016). Relative Citation Ratio (RCR).
  *PLOS Biology*, 14(9), e1002541.
- Waltman, L. et al. (2011). Towards a new crown indicator.
  *Scientometrics*, 87(3), 467--481.
- Uzzi, B. et al. (2013). Atypical Combinations and Scientific Impact.
  *Science*, 342(6157), 468--472.
