---
title: "Reproducibility and Corpus Certificates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducibility and Corpus Certificates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Why certificates?

Every bibliometric analysis starts with a corpus, but most analyses don't
document *how* the corpus was built. scimapR's corpus certificates solve
this: they are structured YAML documents that another researcher can use
to re-derive the exact same corpus.

```{r setup}
library(scimapR)
corpus <- sm_example_corpus(seed = 42)
```

## Provenance tracking

Every record in a scimapR corpus knows where it came from:

```{r provenance}
head(sm_provenance(corpus))
```

## Corpus hashing

```{r hash}
sm_hash_corpus(corpus)
```

## Creating a certificate

```{r certificate}
cert <- sm_certificate(corpus)
str(cert, max.level = 1)
```

## Verifying a certificate

```{r verify}
result <- sm_verify_certificate(corpus, cert)
result$matches
```

## Saving and loading

```{r save-load}
tmp <- tempfile(fileext = ".rds")
sm_save_corpus(corpus, tmp)
loaded <- sm_load_corpus(tmp)
nrow(loaded$works)
```

## Snapshot and diff

```{r snapshot}
snap_path <- tempfile(fileext = ".rds")
sm_snapshot(corpus, snap_path)
```

## Citation

```{r cite}
sm_cite_corpus(corpus)
```

## Locking a corpus

```{r lock}
locked <- sm_lock(corpus, reason = "Final version for publication")
locked$metadata$is_locked

# sm_refresh(locked)  # This would error -- corpus is locked
```
