---
title: "Getting Started with scimapR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with scimapR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/CTTIR/scimapR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CTTIR/scimapR/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/CTTIR/scimapR/actions/workflows/pkgdown.yaml/badge.svg)](https://cttir.github.io/scimapR/)
[![CRAN status](https://www.r-pkg.org/badges/version/scimapR)](https://CRAN.R-project.org/package=scimapR)
[![Codecov test coverage](https://codecov.io/gh/CTTIR/scimapR/branch/main/graph/badge.svg)](https://app.codecov.io/gh/CTTIR/scimapR?branch=main)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/scimapR)](https://cran.r-project.org/package=scimapR)
[![CRAN downloads total](https://cranlogs.r-pkg.org/badges/grand-total/scimapR)](https://cran.r-project.org/package=scimapR)
[![License: MIT](https://img.shields.io/badge/license-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## What is scimapR?

scimapR is a comprehensive toolkit for bibliometric and scientometric
analysis in R. It is designed as a complement to the foundational
[bibliometrix](https://www.bibliometrix.org/) package, adding:

- **Live corpus refresh** with staleness tracking and locking
- **Research questions as first-class objects** with LLM-grounded screening
- **Reproducible-by-construction corpus certificates** for exact re-derivation
- **Author trajectory analysis** with topic pivots and collaborator turnover
- **Equity and representation auditing** with built-in epistemic humility
- **LLM-grounded corpus chat** with retrieval-constrained citations

## Quick start

```{r setup}
library(scimapR)

# Generate a synthetic example corpus
corpus <- sm_example_corpus(n_works = 100, seed = 42)
print(corpus)
```

## Exploring the corpus

```{r works}
# View works
head(corpus$works[, c("work_id", "title", "year", "cited_by_count")])
```

```{r authors}
# View authors
head(corpus$authors[, c("author_id", "display_name")])
```

## Visualisation

All plots use viridis colour palettes by default.

```{r production, fig.cap = "Annual production"}
sm_plot_production(corpus)
```

```{r top-authors, fig.cap = "Top authors"}
sm_plot_top(corpus, level = "authors", n = 10)
```

## Networks and large graphs

Network plots such as `sm_plot_citation_network()` and `sm_plot_collab()`
return `ggraph` objects whose layout is computed lazily at print time. On large
graphs that lazy layout (and the embedded graph) can crash a knitr/`callr`/
workflowr render subprocess. For documents, pass `precompute = TRUE`: the layout
is computed eagerly and a self-contained plain `ggplot` is returned, which you
can `saveRDS()` and print later in the document without recomputing anything ---
even for a graph with a couple of thousand nodes.

The `max_nodes` cap is **opt-in** (`NULL` by default), so existing renders are
never silently downsampled; set it only when you deliberately want to bound a
very large graph (a `cli` message reports when it engages).

```{r network, eval = requireNamespace("ggraph", quietly = TRUE)}
p <- sm_plot_citation_network(corpus, precompute = TRUE)
# saveRDS(p, "fig_network.rds")  # print later in the document, cheaply
p
```

## Filtering

```{r filter}
recent <- sm_filter_works(corpus, year_range = c(2020, 2024))
nrow(recent$works)
```

## Next steps

- See `vignette("ingestion")` for building corpora from real data
- See `vignette("relationship-to-bibliometrix")` for interop with bibliometrix
- See `vignette("embeddings-and-clusters")` for semantic analysis
- Run `sm_run_app()` for the interactive Shiny explorer

# Use of LLM tools

Portions of this package were prepared with assistance from large language model tooling for
narrowly defined, non-authorial tasks: copyediting, prose smoothing, Markdown/LaTeX formatting,
scaffolding of boilerplate files (CI configs, build scripts), code refactoring. The tools used were [Chat AI](https://kisski.gwdg.de/leistungen/2-02-llm-service/),
the LLM service of KISSKI (GWDG), and a self-hosted **Mistral Small (24B, Apache-2.0)** run locally via
[Ollama](https://ollama.com/) and the `ollamar` R package — local inference only, with no data sent to
third parties for the self-hosted model.

All scientific claims, methodological choices, analyses, interpretations, and conclusions are the
author's own. No LLM-generated text was incorporated without review and revision, and every reference
was verified against its DOI, arXiv ID, or ISBN.
