---
title: "Getting Started with scimapR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with scimapR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## What is scimapR?

scimapR is a comprehensive toolkit for bibliometric and scientometric
analysis in R. It is designed as a complement to the foundational
[bibliometrix](https://www.bibliometrix.org/) package, adding:

- **Live corpus refresh** with staleness tracking and locking
- **Research questions as first-class objects** with LLM-grounded screening
- **Reproducible-by-construction corpus certificates** for exact re-derivation
- **Author trajectory analysis** with topic pivots and collaborator turnover
- **Equity and representation auditing** with built-in epistemic humility
- **LLM-grounded corpus chat** with retrieval-constrained citations

## Quick start

```{r setup}
library(scimapR)

# Generate a synthetic example corpus
corpus <- sm_example_corpus(n_works = 100, seed = 42)
print(corpus)
```

## Exploring the corpus

```{r works}
# View works
head(corpus$works[, c("work_id", "title", "year", "cited_by_count")])
```

```{r authors}
# View authors
head(corpus$authors[, c("author_id", "display_name")])
```

## Visualisation

All plots use viridis colour palettes by default.

```{r production, fig.cap = "Annual production"}
sm_plot_production(corpus)
```

```{r top-authors, fig.cap = "Top authors"}
sm_plot_top(corpus, level = "authors", n = 10)
```

## Filtering

```{r filter}
recent <- sm_filter_works(corpus, year_range = c(2020, 2024))
nrow(recent$works)
```

## Next steps

- See `vignette("ingestion")` for building corpora from real data
- See `vignette("relationship-to-bibliometrix")` for interop with bibliometrix
- See `vignette("embeddings-and-clusters")` for semantic analysis
- Run `sm_run_app()` for the interactive Shiny explorer
